Data-Cleaning.Rmd
CSE-160 Final-Project
Paolo Bartolucci
5/5/2024

Used to load core df used for train/test

[Dataset2](https://data.cityofnewyork.us/Public-Safety/Motor-Vehicle-Collisions-Crashes/h9gi-nx95/about_data)

This is "DataSet2" From the original Report. Dataset 1 contained lat and long data that I was unable to convert to zip code.

This is crash data in NYC. The original dataset is around ~2 mil rows, I used the online filter engine to only include crashes from 2024. Converted the dataset to around 250k rows. There is enough data within the rows to make a difference between fatal vs nonfatal crashes, but that can be changed in the future. For now I will aggregate just a sum of total crashes.

The Result of this is a vector with accident frequencies by zipcode


```{r}


crash_data <- read.csv("data/Motor_Vehicle_collisions.csv")

crash_data<- na.omit(crash_data)

crash_data <- crash_data$ZIP.CODE

crash_data <- as.data.frame(table(crash_data))

names(crash_data) <- c("zipcode","freq")


```

[Dataset3](https://datacommons.techsoup.org/tools/download#pt=CensusZipCodeTabulationArea&place=geoId%2F3651000&sv=Median_Income_Household&dtType=ALL&facets=%7B%7D)

This dataset is the aggregation of existing Cesus data that groups data directly into NYC zip Codes. This contains the median incomes of families by zipcode from 2014-2020.

The income_data df contains the median incomes of NYC zipcodes for 2020.


```{r}
income_data <- read.csv("data/NYC_Census_Income.csv")

income_data <- income_data[,c(2,3,4)]
income_data <- income_data[income_data$Date.Median_Income_Household == 2020,]
income_data <- income_data[,c(1,3)]
names(income_data) <- c("zipcode","median_income")

data <- merge(crash_data, income_data, by="zipcode")
```



[Dataset4]()

check if lat long exists in shapefile


```{r}
rental_data <- read.csv("data/Aggregated_Rental.csv")

rental_target <- rental_data[,c('price','bedrooms','zipcode')]


data<- merge(data, rental_target, by="zipcode")
```


